4.2: Difference-in-Difference Models

### 2019-12-05 static/slides/23-slides.Rmd
Final tweak for slides 23

```{r, fig.retina=3, fig.align="center", fig.width=13}
ggplot(data = tibble(x=c(0,10),
                     y=c(0,10)))+
  aes(x = x,
      y = y)+
  geom_label(x=1,y=2, label="Correlation", size =10, color = "#fde0dd")+
  geom_label(x=5,y=2, label="Causation", size = 10, color = "#7a0177")+

@@

Again, this toolkit of research designs to identify causal effects is the economist's comparative advantage that firms and governments want!

@@

- Regression equation:
$$\widehat{Enrolled_{it}} = \beta_0+\beta_1 \, Georgia_{it}+\beta_2 \,  After_{it}+\beta_3 \, (Georgia_{it} \times After_{it})$$

@@

.font70[
```{r, echo=T}
DND_reg<-lm(InCollege ~ Georgia + After + Georgia:After, data = hope)
summary(DND_reg)
```
]

@@

# group mean for Georgian before 1992
hope %>%
  filter(Georgia==1,
         After==0) %>%
  summarize(prob = mean(InCollege))
```
]

@@

# group mean for Georgian AFTER 1992
hope %>%
  filter(Georgia==1,
         After==1) %>%
  summarize(prob = mean(InCollege))
```
]

@@

|    | Neighbors | Georgia | **Group Diff** $(\Delta Y_i)$ |
|----|---------|-----------|---------------------------|
| Before | $0.406$ | $0.301$ | $-0.105$ |
| After | $0.402$ | $0.386$ | $0.016$ |
| **Time Diff** $(\Delta Y_t)$ | $-0.004$ | $0.085$ | **Diff-in-diff** $0.089$ |

@@

- DND can be **generalized** with a **two-way fixed effects** model:  	
$$\widehat{Y_{it}}=\alpha_i+\theta_t+\beta_3 (\text{Treated}_i * \text{After}_{it})+\nu_{it}$$
    - $\alpha_i$: **group fixed effects** (treatments/control groups)
    - $\theta_t$: **time fixed effects** (pre/post treatment)

@@

- Can also add control variables that vary within units and over time 
$$\widehat{Y_{it}}=\alpha_i+\theta_t+\beta_3 \, (\text{Treated}_i \times \text{After}_{it})+\beta_4 X_{it}+\nu_{it}$$

>>>>

Again, **this toolkit** of research designs to **identify causal effects** is the economist's **comparative advantage** that firms and governments want!

```{r, fig.retina=3, fig.align="center", fig.width=12}
ggplot(data = tibble(x=c(0,10),
                     y=c(0,10)))+
  aes(x = x,
      y = y)+
  geom_label(x=1,y=2, label="Correlation", size =10, color = "#fde0dd")+
  geom_label(x=5,y=2, label="Causation", size = 10, color = "#7a0177")+

@@

@@

- Regression equation:
$$\widehat{Enrolled_{it}} = \beta_0+\beta_1 \, Georgia_{i}+\beta_2 \,  After_{t}+\beta_3 \, (Georgia_{i} \times After_{t})$$

@@

.font70[
```{r, echo=T}
DND_reg<-lm(InCollege ~ Georgia + After + Georgia:After, data = hope)
summary(DND_reg)
```
]
]

@@

```
]

@@

```
]

# Example: Comparing Group Means in R II

.pull-left[
```{r, echo=T}
# group mean for Georgian before 1992
hope %>%
  filter(Georgia==1,
         After==0) %>%
  summarize(prob = mean(InCollege))
```
]

.pull-right[
```{r, echo=T}
# group mean for Georgian AFTER 1992
hope %>%
  filter(Georgia==1,
         After==1) %>%
  summarize(prob = mean(InCollege))
```
]

@@

|    | Neighbors | Georgia | **Group Diff** $(\Delta Y_i)$ |
|----|--------:|----------:|--------------------------:|
| Before | $0.406$ | $0.301$ | $-0.105$ |
| After | $0.402$ | $0.386$ | $0.016$ |
| **Time Diff** $(\Delta Y_t)$ | $-0.004$ | $0.085$ | **Diff-in-diff**: $0.089$ |

@@

- DND can be **generalized** with a **two-way fixed effects** model:  	
$$\widehat{Y_{it}}=\alpha_i+\theta_t+\beta_3 (\text{Treated}_i * \text{After}_{t})+\nu_{it}$$
    - $\alpha_i$: **group fixed effects** (treatments/control groups)
    - $\theta_t$: **time fixed effects** (pre/post treatment)

@@

- Can also add control variables that vary within units and over time 
$$\widehat{Y_{it}}=\alpha_i+\theta_t+\beta_3 \, (\text{Treated}_i \times \text{After}_{t})+\beta_4 X_{it}+\nu_{it}$$
